---
title: "para_est"
author: "Yu He"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(deSolve)
library(R0)
```


# Part 1, import the real influenza data
```{r}
# Define the reference date
reference_date <- as.Date("2020-01-01")

para_R0_est = read_csv("dat0.csv") %>% 
  subset(date >= 43830) %>%
  janitor::clean_names() %>% 
  mutate(
    date = reference_date + (date - 43835)
  ) %>% 
  select(date, real_flu_case)


## Always plot and check the data first
par(mar = c(3, 3, 1, 1), cex = 1.2, mgp = c(1.5, .5,0))
plot(para_R0_est$date, para_R0_est$real_flu_case, xlab ='time', ylab = 'Cases', pch = 20)

para_R0_est$real_flu_case
```

## modify the orginal data and insert dates for everyday
```{r}
# create the original data frame
dates <- c("2020-01-01", "2020-01-08", "2020-01-15", "2020-01-22", "2020-01-29", "2020-02-05", "2020-02-12", "2020-02-19", "2020-02-26", "2020-03-04", "2020-03-11", "2020-03-18")
real_flu_cases <- c(591, 1211, 1159, 1181, 1003, 320, 93, 23, 18, 7, 17, 25)
df <- data.frame(date = as.Date(dates), real_flu_case = real_flu_cases)

# create a new data frame with every day between min and max date
min_date <- min(df$date)
max_date <- max(df$date)
date_range <- seq(min_date, max_date, by='day')
new_df <- data.frame(date = date_range)

# merge the new data frame with the original data frame
df <- merge(x=new_df, y=df, by='date', all.x=TRUE)
df$real_flu_case[is.na(df$real_flu_case)] <- 0


# plot the real flu cases over time
ggplot(df, aes(x=date, y=real_flu_case)) +
  geom_line() +
  labs(x="Date", y="Real Flu Cases", title="Real Flu Cases over Time")
```


# Part 2, alculate the cumulative incidence using the daily incidence dataset and plot the log(cumulative incidence) v. time
```{r}
# cumpute the cumulative incidence using the cumsum function
cumI = cumsum(df[,2]);

# calculate the log - cumulative incidence
log_cumI = log(cumsum(df[,2]))

par(mar = c(3, 3, 1, 1), cex = 1.2, mgp = c(1.5, .5,0))
plot(df$date, log_cumI, main = "Log - Cumulative incidence for 1918 Germany flu", xlab ='Date', ylab = 'log - cumulative incidence', pch = 20)


```


# Part 3, Assume a generation time of 3.5 days, estimate R0 from the exponential growth phase of the epidemic (try the first 12, 34, and 50 days), using the following equation:
# Report your R0 estimates for the three time periods (i.e., the first 7, 14, and 21 days)
```{r}
# EXAMPLE: FOR THE FIRST 11 DAYS
# Fit to the data during the first 11 days:
D = 3.5; # set the generation time to 3.5 days
Ndays = 12;  # ADJUST THE NUMBER OF DAYS INCLUDED IN THE FIT HERE
tm1 = 1:Ndays;
fit1 = lm(log(cumI[1:Ndays]) ~ tm1)
summary(fit1)

# compute R0 based on the model-fit
R = 1 + fit1$coefficients[2]*D   # slope: fit1$coefficients[2]
# R0 = 1.48
```


```{r}
# EXAMPLE: FOR THE FIRST 34 DAYS
# Fit to the data during the first 11 days:
D = 3.5; # set the generation time to 3.5 days
Ndays = 34;  # ADJUST THE NUMBER OF DAYS INCLUDED IN THE FIT HERE
tm1 = 1:Ndays;
fit1 = lm(log(cumI[1:Ndays]) ~ tm1)
summary(fit1)

# compute R0 based on the model-fit
R = 1 + fit1$coefficients[2]*D   # slope: fit1$coefficients[2]
# R0 = 1.26
```


```{r}
# EXAMPLE: FOR THE FIRST 50 DAYS
# Fit to the data during the first 11 days:
D = 3.5; # set the generation time to 3.5 days
Ndays = 50;  # ADJUST THE NUMBER OF DAYS INCLUDED IN THE FIT HERE
tm1 = 1:Ndays;
fit1 = lm(log(cumI[1:Ndays]) ~ tm1)
summary(fit1)

# compute R0 based on the model-fit
R = 1 + fit1$coefficients[2]*D   # slope: fit1$coefficients[2]

# R0 = 1.16
```


```{r}
# EXAMPLE: FOR THE FIRST 50 DAYS
# Fit to the data during the first 11 days:
D = 3.5; # set the generation time to 3.5 days
Ndays = 50;  # ADJUST THE NUMBER OF DAYS INCLUDED IN THE FIT HERE
tm1 = 1:Ndays;
fit1 = lm(log(cumI[1:Ndays]) ~ tm1)
summary(fit1)

# compute R0 based on the model-fit
R = 1 + fit1$coefficients[2]*D   # slope: fit1$coefficients[2]
```












